---
title: "ANOVA"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'doc/', warning=FALSE)
```

## Import data

ChickWeight data describes a mixed design: a group of 45 chickens were split into 4 groups, each group was given a different diet. Each chick was weighed at 6 time points (4 week intervals).  
The sav file has the following variables:

- Chick - the chich ID
- Diet - 1-4 the Diet number, with 1 being the standard diet.
- T0:T20 - weigts at each of the 6 time points.

```{r}
library(haven)

# import wide format
df_wide <- read_spss('data/ChickWeight.sav')

head(df_wide)
```

We need the data in the long format:

```{r, message=FALSE}
library(tidyverse)

df_long <- df_wide %>% 
  gather("Time","weight",T0:T20)


head(df_long)
```

## repeated measure ANOVA

We will fit a mixed-anova model with `afex`:

```{r, message=FALSE}
library(afex)

# fit using `aov_ez`
fit <- aov_ez(id = "Chick", dv = "weight", data = df_long,
              between = c("Diet"), within = "Time")
fit
```

We can change some of the defults, such as $\eta^2_2$ for effect-sizes, and not correcting for GG:

```{r, message=FALSE}
fit <- aov_ez(id = "Chick", dv = "weight", data = df_long,
              between = "Diet", within = "Time",
              anova_table = list(es = "pes", correction = "none"))
fit
```

## Contrasts

Before running contrasts in `emmeans`, we need to set some of `afex` options, to ensure that MSE and DF for between-subject contrasts are computed/estimated using the multivariate method (producing the same results as SPSS ans Statistica):

```{r}
afex_options(emmeans_model = "multivariate")
```

`emmeans` gives you much control over how contrasts are calculated.  
But... with great power comes great responsibility... (This is probably true for most anything in R)

### simple effects

```{r}
library(emmeans)  

joint_tests(fit, by = 'Diet')
joint_tests(fit, by = 'Time')
```


### Simple plot (at a glance)

```{r}
emmip(fit, ~Time, CIs = TRUE)
```

### compute contrasts with `contrast`

We first look create a refernce grid (which we can print to look at the predicted means):

```{r}
em_time <- emmeans(fit, ~Time)
em_time
```

Using some methods that come baked into `emmeans`:
```{r}
contrast(em_time, "poly")
contrast(em_time, "consec") # using built in methods, emmeans knows when contrasts are dependant
contrast(em_time, "consec", adjust = "fdr") # or adjust = "tukey"
contrast(em_time, "consec", adjust = "none")


```

We can also make custom methods. I will demonstrate here a simple method, by you can [read more about creating more advanced methods](https://cran.r-project.org/web/packages/emmeans/vignettes/comparisons.html#linfcns):
```{r}
time_contrasts.emmc <- function(...){
  data.frame(
    half_time = c(-1,-1,-1,1,1,1)/3, # divide to get un-inflated estimate (has no effect on testing)
    start_end = c(-1,0,0,0,0,1)
  )
}

contrast(em_time, "time_contrasts")

```

## Interactions
We can do the same for interactions:

```{r}
emmip(fit, Diet~Time, CIs = T)

em_interaction <- emmeans(fit, ~ Diet * Time)
em_interaction


contrast(em_interaction, "consec", by = 'Diet')
contrast(em_interaction, "pairwise", by = 'Time')
contrast(em_interaction, "pairwise", by = 'Time', adjust = 'none')
contrast(em_interaction, interaction = c("pairwise","consec"))

# or... using custom contrasts:

diet_contrasts.emmc <- function(...){
  data.frame(
    none_all = c(3,-1,-1,-1)/3
  )
}

contrast(em_interaction, interaction = c('diet_contrasts','time_contrasts'))
```

